---
title: "Mini Data Analysis"
author: "William Laplante"
date: "05/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE,message=FALSE}
library(datateachr)
library(tidyverse)
library(knitr)
library(DT)
```

## Task 1 : Choosing favorite dataset

Out of the 7 datasets, we pick the following:<br>
1 : apt_buildings<br>
2 : cancer_sample<br>
3 : flow_sample<br>
4 : vancouver_trees<br>

Now, we explore these datasets to find out which one we would like to analyze. We look at different attributes of the data, such as the dimensions of the datasets, the variables and their datatypes, and the basic statistics describing our data.<br>

#### Exploring the data


##### Variables, data types and dimensions
```{r}
glimpse(cancer_sample, fillContainer = T)
```

```{r}
glimpse(apt_buildings)
```

```{r}
glimpse(flow_sample)
```

```{r}
glimpse(vancouver_trees)
```

##### Basic statistics

```{r}
summary(cancer_sample)
```

```{r}
summary(apt_buildings)
```

```{r}
summary(flow_sample)
```

```{r}
summary(vancouver_trees)
```

For our analysis, we decide to use the cancer_sample and the vancouver_trees datasets. These datasets were picked because they fall under the biostatistics field, and thus they fit my interests.<br>

We then choose 1 research question that we would like to answer with each dataaset : <br>
cancer_sample : How does the area of a cancer tumor affects the diagnosis? <br>
vancouver_trees : How does the geographic position of a tree affects the volume of a tree, by species ?<br>

From these questions, the cancer_sample one is the most interesting one. Hence we choose the cancer_sample dataset for our analysis.

## Task 2 : Exploring dataset

1) We plot the distribution of tumor area by diagnosis, where B means benign and M means malignant. The reason we do this is to observe if the area of a malignant cancer tends to be bigger in comparison with a benign cancer. 
```{r}
ggplot(cancer_sample, aes(x=area_mean)) + geom_histogram(aes(color = diagnosis))
```

2) We explore the relation between smoothness and compactness by type of diagnosis. We do this to explore if smoothness (local variation in radius lengths) is correlated to compactness (perimeter² / area — 1.0).
```{r}
ggplot(cancer_sample, aes(x=smoothness_mean, y=compactness_mean)) + geom_point(aes(color=diagnosis))
```
  
We do a boxplot of concavity by diagnosis. This is to explore whether the median concavity is bigger for malignant cancers compared to benign cancers. 
```{r}
ggplot(cancer_sample, aes(x=diagnosis, y=concavity_mean)) + 
    geom_boxplot()
```

We look at the density plot of cancer symmetry by diagnosis. This will help us explore whether benign cancers have more symmetry than malignant cancer.
```{r}
ggplot(cancer_sample, aes(symmetry_mean)) +
    geom_density(aes(color=diagnosis))
```

## Task 3 : Research questions

We now establish 4 research questions based on our exploration of the data.

1) How does the area of a cancer affects the diagnosis?
2) How strongly do smoothness and compactness correlate, and is it dependent on diagnosis?
3) What is the variable in this dataset that is the best indicator of a malignant cancer?
4) Is it harder to perform radius measurements on malignant cancers compared to benign cancers? That is, is the standard error in radius measurements bigger for malignant cancers?


