---
title: "Mini Data Analysis"
author: "William Laplante"
date: "05/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE,message=FALSE}
library(datateachr)
library(tidyverse)
library(knitr)
library(DT)
```

## Task 1 : Choosing favorite dataset

Out of the 7 datasets, we pick the following:<br>
1 : apt_buildings<br>
2 : cancer_sample<br>
3 : flow_sample<br>
4 : vancouver_trees<br>

Now, we explore these datasets to find out which one we would like to analyze. We look at different attributes of the data, such as the dimensions of the datasets, the variables and their datatypes, and the basic statistics describing our data.<br>

#### Exploring the data


##### Variables, data types and dimensions
```{r}
glimpse(cancer_sample, fillContainer = T)
```

```{r}
glimpse(apt_buildings)
```

```{r}
glimpse(flow_sample)
```

```{r}
glimpse(vancouver_trees)
```

##### Basic statistics

```{r}
summary(cancer_sample)
```

```{r}
summary(apt_buildings)
```

```{r}
summary(flow_sample)
```

```{r}
summary(vancouver_trees)
```

For our analysis, we decide to use the cancer_sample and the vancouver_trees datasets. These datasets were picked because they fall under the biostatistics field, and thus they fit my interests.<br>

We then choose 1 research question that we would like to answer with each dataaset : <br>
cancer_sample : How does the area of a cancer tumor affects the diagnosis? <br>
vancouver_trees : How does the geographic position of a tree affects the volume of a tree, by species ?<br>

From these questions, the vancouver_trees one seems the most interesting one. Hence we choose the vancouver_trees dataset for our analysis.

## Task 2 : Exploring dataset

1) We plot the distribution of tree diameter.
```{r}
ggplot(vancouver_trees, aes(x=diameter)) + geom_histogram()
```

2) We create a new variable called vol_estimate, which has the goal to estimate the volume of each tree. We assume a cylindral shape for our estimate. Thus, the formula should be:
$$
V = \pi (\frac{D}{2})^{2} \times H,
$$
where $D$ and $H$ are the tree's diameter and height respectively. Since the variable height_range_id is following a convention such that the value is of 0-10 for every 10 feet, we first add 1 to each value and multiply by 10. This will give us our height column for the volume calculation.


```{r}
transmute(vancouver_trees, vol_estimate = 10*(height_range_id+1) * pi * (diameter/2)^2)
```

```{r}
datatable(head(vancouver_trees))
```


