---
title: "Mini-data-analysis-m2"
author: "William Laplante"
date: "12/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(datateachr)
library(plyr)
library(tidyverse)
library(DiscriMiner)
library(corrr)
```

## Task 1 : Process and summarize your data

### 1.1 : Research questions

1) How does the area of a cancer affects the diagnosis?

2) How strongly do smoothness and compactness correlate, and is it dependent on diagnosis?

3) What is the variable in this dataset that is the best indicator of a malignant cancer?

4) Is it harder to perform radius measurements on malignant cancers compared to benign cancers? That is, is the standard error in radius measurements bigger for malignant cancers?

### 1.2 : Summarizing and Graphing 

#### Research question 1

We compute the summary statistics of the area of a cancer for malignant cancers and for bening cancers. This gives us insights on the spread of our data and where it is centered. Also, we can compare the mean and median to see if there are significant outliers. 

##### Malignant cancer - summary statistics
```{r}
cancer_sample %>% filter(diagnosis == "M") %>% select(area_mean) %>% summary()
```
##### Benign cancer - summary statistics
```{r}
sum_B = cancer_sample %>% filter(diagnosis == "B") %>% select(area_mean) %>% summary()
sum_B
```
We also compute and store the mean of area_mean for each of the two diagnosis. This will be used later for plotting.
```{r}
mu = ddply(cancer_sample, "diagnosis", summarise, grp.mean=mean(area_mean))
mu
```
##### Graphing
Now, we can visualize the existing difference in summary statistics between the benign cancer and malignant cancer by making a histogram for both cases. We play on the transparency to show the small overlap between the two histograms.
```{r}
p <- ggplot(cancer_sample, aes(x=area_mean, fill=diagnosis, color=diagnosis)) + geom_histogram(position="identity", alpha=0.5)
p +   geom_vline(data=mu, aes(xintercept=grp.mean, color=diagnosis), linetype="dashed")
```

#### Research question 2

We look into the summary statistics of smoothness and compactness. Also, we look at the covariance and correlation between the two.

##### smoothness - summary statistics
```{r}
cancer_sample %>% select(smoothness_mean) %>% summary()
```
##### compactness - summary statistics
```{r}
cancer_sample %>% select(compactness_mean) %>% summary()
```

##### Correlation between the variables
```{r}
cor(cancer_sample$smoothness_mean, cancer_sample$compactness_mean)
```

##### Graphing
Now, we make a plot illustrating the correlation between the two variables. We add a fitted regression line to show the correlation. Note that for now, we ignore the diagnosis component of the data.

```{r}
ggplot(cancer_sample, aes(smoothness_mean, compactness_mean)) + geom_point() + geom_smooth(method=lm)
```

#### Research question 3

We look at the correlation between all numerical variables in the dataset and the diagnosis categorical variable. To do so, we convert the diagnosis column to a numerical datatype, so that malignant ("M") becomes 1 and benign ("B") becomes 0. This makes it so we can compute the correlation matrix.
```{r}
cancer_sample_numeric <- cancer_sample %>% replace(cancer_sample=="M","1") %>% replace(cancer_sample=="B","0") %>% transform(diagnosis=as.numeric(diagnosis))
cancer_sample_cormatrix = cor(cancer_sample_numeric)
```

```{r}
cormatrix_df <- as_cordf(cancer_sample_cormatrix)
cormatrix_df
```
We can then split the correlation coefficients into 3 categories : High, Medium and Low. This allows us to understand the influence of the existing variables on the cancer diagnosis. We illustrate this splitting using a bar graph.
```{r}

category_corr <- cormatrix_df %>% mutate(category=cut(diagnosis, breaks=c(-Inf,0.33, 0.66, Inf), labels=c("Low", "Medium", "High"))) %>% select(term, diagnosis, category) %>% filter(!is.na(diagnosis)) %>% rename(variable=term, corr_with_diagnosis=diagnosis)
category_corr
```
Here, the graph shows the number of variables falling under each category for the correlation coefficient with the cancer diagnosis.
```{r}
ggplot(category_corr, aes(x=category)) + geom_bar()
```

#### Research question 4

For this research question, we can create a new variable, which would be the absolute difference between the mean and the worst measurement for each diagnosis. If the worst is far from the mean, it tells us that there is a high variance between measurements of the same variable. 

```{r}
data_mean <- cancer_sample %>% select(radius_mean : fractal_dimension_mean) 

data_worse <- cancer_sample %>% select(radius_worst : fractal_dimension_worst)

radius_diff = rename_with(abs(data_mean-data_worse), ~gsub("_mean", "_diff", .x, fixed=TRUE)) %>% select(radius_diff)
head(radius_diff)
```
We can then compute some statistics to see how variable radius measurements are :
```{r}
summary(radius_diff$radius_diff)
```
```{r}
summary(data_mean$radius_mean)
```

Doing so gives us input on the range of radius measurements, i.e. how precise the given values for radius_mean are.We can visualize this by using a boxplot :

```{r}
radius_df = data.frame(cancer_sample$radius_mean, radius_diff)
ggplot(stack(radius_df), aes(x=ind, y=values)) + geom_boxplot()
```


